### 2009-06-15 paper.tex
paper updates

We first consider the {\em perfect alignment rate}; that is, the number of times each method gets the alignment exactly correct.
Theory predicts that ML inference, using the correct model and parameters, should maximize the perfect alignment rate.
Inspecting \figref{perfect}, we find this to be the case, except when the outgroup is very distant
and the best perfect alignment rate drops below 0.05 (at which point 125 trials may not be sampling it sufficiently).
We also note that the ML version of \stemloc\ is near-optimal, and slightly better than the posterior-decoding \stemloc-AMA.
Finally, we note that the structure-blind models (TKF91 and Long Indel) perform consistently worse than the structure-aware methods,
with no decisive advantage shown by either TKF91 (linear gap penalty) or Long Indel (affine gap penalty).

@@

% figure produced automatically with simulations/plots.r from simulations/inferred/indiegram.analysis
\begin{figure}[!ht]
  \centering
  % \includegraphics[width=5in]{figs_plos/perfect.pdf}
  \caption{
    \textbf{Dependence of perfect alignment rate on outgroup branch
      length.}
    We investigated the perfect alignment rate
    (the number of times that a given alignment program estimates the alignment 100\% correctly, with no errors)
    as a function of outgroup branch length, simulating the
    evolution of three structural RNAs under our model.  The two
    sister species were equidistant from the ancestral sequence, with
    a branch length of 1, corresponding to one expected
    substitution per site in loop sequence.  We allowed the outgroup
    branch length $t$ to vary between $[0, 2.5]$.  We selected 25 equally spaced values of $t$ in this range, spaced $0.1$ apart,
    and simulated 25 alignments for each value of $t$, using TKFST model parameters described in the text.
    Since the perfect alignment rate is rather low, we further aggregated the $t$-values for this plot into bins of five;
    thus, for example, the bin named ``$0 to 0.5$'' includes $t \in \{ 0, 0.1, 0.2, 0.3, 0.4 \}$.
    The vertical axis counts the number of perfect alignments in each bin, out of a total of 125 trials per bin.
}
  \figlabel{perfect}
\end{figure}

>>>>

We first consider the {\em perfect alignment rate}; that is, the number of times each method gets the alignment exactly correct.
Theory predicts that ML inference, using the correct model and parameters, should maximize the perfect alignment rate.
Inspecting \figref{perfect}, we find this to be the case, except where the outgroup is very distant and the bins are probably undersampled
(when the perfect alignment rate drops below $\sim 5/125$, 125 trials may be too few).
We also note that the ML version of \stemloc\ is near-optimal, and slightly better than the posterior-decoding \stemloc-AMA.
Finally, we note that the structure-blind models (TKF91 and Long Indel) perform consistently worse than the structure-aware methods,
with no decisive advantage shown by either TKF91 (linear gap penalty) or Long Indel (affine gap penalty).

@@

% figure produced automatically with simulations/plots.r from simulations/inferred/indiegram.analysis
\begin{figure}[!ht]
  \centering
  % \includegraphics[width=5in]{figs_plos/perfect.pdf}
  \caption{
    \textbf{Dependence of perfect alignment rate on outgroup branch
      length.}
    We investigated the perfect alignment rate
    (the number of times that a given alignment program estimates the alignment 100\% correctly, with no errors)
    as a function of outgroup branch length, simulating the
    evolution of three structural RNAs under our model.  The two
    sister species were equidistant from the ancestral sequence, with
    a branch length of 1, corresponding to one expected
    substitution per site in loop sequence.  We allowed the outgroup
    branch length $t$ to vary between $[0, 2.5]$.  We selected 25 equally spaced values of $t$ in this range, spaced $0.1$ apart,
    and simulated 25 alignments for each value of $t$, using TKFST model parameters described in the text.
    Since the perfect alignment rate is rather low, we further aggregated the $t$-values for this plot into bins of five;
    thus, for example, the bin named ``$0 to 0.4$'' includes $t \in \{ 0, 0.1, 0.2, 0.3, 0.4 \}$.
    The vertical axis counts the number of perfect alignments in each bin, out of a total of 125 trials per bin.
}
  \figlabel{perfect}
\end{figure}
